SAT Solving for Vesicle Traffic Systems in Cells

### 2017-06-06 paper/main.tex
updated-paper-fixed-a-bug
paper folder organized

\begin{document}
\vspace*{0.2in}

% Title must be 250 characters or less.
\begin{flushleft}
{\Large
\textbf\newline{ Vesicle traffic system and constraints.} % Please use 
}
\newline
\\
\end{flushleft}

\section{Formula variables}
Basic notation for the formula we are trying to generate.
\begin{enumerate}
\item edge = e(i,j,q)
      There is an qth edge between node i and j.

\item dump = d(i,j,q)
       drop edge(i,j,q) from the graph.

\item presence node = n(i,k) 
	  kth molecule is present on the node i.

\item active\_node = a(i,k) 
      kth molecule is active on the node i.

\item presence\_edge = e(i,j,q,k)
      There is an qth edge between node i and j with kth molecule present.

\item active\_edge = b(i,j,q,k)
      There is an qth edge between node i and j with kth molecule active.

\item reachability = r(i,j,k,z)
      Node i and j are reachable with kth molecule present in z steps.

\item pairing matrix = p(k,k')
      Molecule k and k' are pairing molecules in the matrix.

\item Type of the Boolean function = sorts (takes k-1 arguments Bool and return Bool) 
\item Boolean function on nodes Function (fn\_{m},*sorts)

\item Boolean function on edge Function(fe\_{m},*sorts)

\end{enumerate}

\section{Regulation on nodes and edges}
Regulation of a model is based on the kind of the model under consideration. for example model 1 (-V 1) has two different arbitrary Boolean function dictating the activity of the molecules on edge and nodes and for variation 3 (-V 3) activity on edge is driven by SNARE-SNARE inhibition.

We have build constraints A0, A1 for these two regulations. \newline

A0: No regulation on nodes.
\[ \bigwedge\limits_{i,k} n_{i,k} == a_{i,k} \, \]

A1: No regulation on edges.
\[ \bigwedge\limits_{i,j,q,k} e_{i,j,q,k} == b_{i,j,q,k} \, \]

A0: Boolean regulation on nodes. Activity of a molecule k on a node is defined as a Boolean function of presence of other molecule present on that node.
\[ \bigwedge\limits_{i,k} n_{i,k} \supset a_{i,k} ==   f_n[k] (\bigvee_{ k' != k} n_{i,k'}) \, \]

A1: Boolean regulation on edges. Activity of a molecule k on a node is defined as a Boolean function of presence of other molecule present on that node.
\[ \bigwedge\limits_{i,j,q,k} e_{i,j,q,k} \supset b_{i,j,q,k} == f_e[k](\bigvee_{k' != k} e(i,j,q,k')) \, \]

A1: SNARE-SNARE inhibition. Inhibition of the edges are driven by the pairing matrix.
\[ \bigwedge\limits_{i,j,i != j, q, k} (\bigwedge_{k' != k} p_{k,k'} \supset e_{i,j,q,k'}) \supset \neg b(i,j,q,k)) \, \]

\[ \bigwedge\limits_{i,j,i != j, q, k} \neg [(\bigwedge_{k' != k} p_{k,k'} \supset e_{i,j,q,k'})] \supset b(i,j,q,k)) \, \]

\section{Basic constraints on the edges and nodes}
The edge labels are subset of the node label of source and target compartment. \newline

C0: The edge labels are subset of the node label of source compartment.

\[ \bigwedge\limits_{i,j,k} e_{i,j,k} \supset a_{i,k} \, \]

C1: The edge labels are subset of the node label of target compartment.

\[ \bigwedge\limits_{i,q}  e_{i,j,k} \supset a_{j,k} \, \]  
%C1 =  e_ijk -> aik and e_ijk -> ajk

C2: Self edges are not allowed.

\[  \bigwedge\limits_{i,q} \neg e_{i,i,q} \, \]

C4: Condition on p\_kk'. Diagonal blocks should be all 0's.

\[ \bigwedge\limits_{(x < M/2 \, \land  \, y < M/2) \lor  (x >= M/2 \land y >= M/2)} \neg p(x,y) \, \]

C5: Activity on the node. A molecule should be present to be active on a node.  
\[ \bigwedge\limits_{i,k} a_{i,k} \supset n_{j,k} \, \]

% - F0: b(i,j,k) $\supset$ e(i,j,k)

\section{Main constraints}
Main constraints to be followed. \newline

F0:  An edge has to have one present molecule. F0 looks redundant.

\[ \bigwedge\limits_{i,j,q} \bigvee_k e_{i,j,q,k} \supset e_{i,j,q} \, \]

F1: If molecule is active on an edge then it should be present on the edge.

\[ \bigwedge\limits_{i,j,q,k} a_{i,j,q,k} \supset e_{i,j,q,k}\, \]

\section{Steady state specification}
Each molecule leaving the node on a vesicle should come back to its source node in a cycle, i.e., \textbf{for every molecule leaving the node there exists a cycle with that molecule present on \textit{each of the edges and nodes} of the path taken}.

\subsection{Reachability definition and stability condition}
If molecule is active on an edge then it should be present on the edge. \newline

F3: stability condition. Source node is reachable by target node with that molecule present in p steps.

% I have implemented: Changed the implementation in code too.
%\[ \bigwedge\limits_{i,j,q,k} e_{i,j,q,k} \supset (r_{j,i,k,0}  \lor r_{j,i,k,1} \lor ... \lor r_{j,i,k,p}) \, \]

% Update: Better 
\[ \bigwedge\limits_{i,j,k} (\bigvee_{q} e_{i,j,q,k}) \supset (r_{j,i,k,0}  \lor r_{j,i,k,1} \lor ... \lor r_{j,i,k,p}) \, \]

F2: Reachability definition.
\[ \bigwedge\limits_{i,j,k,p} r_{i,j,k,p} \supset (\bigvee_{q} \, e_{i,j,q,k} \lor \bigvee_{i\neq i^{\prime}} ( \, \bigvee_{q} e_{i,i^{\prime},q,k}) \land r_{i^{\prime},j,k,p-1} ) \, \]

\section{ Fusion rule Formula}

Fusion rules consist of two different mechanisms.

\begin{enumerate}
\item  A \textbf{SNARE pairing mechanism} which determines compatible Q-R pairs on vesicles and compartments that can cause fusion.
\item \textbf{Regulatory mechanisms} on edges and on nodes (possibly distinct) which regulate SNARE activity based on the presence/absence of other molecules on the corresponding node or edge.
\end{enumerate}

F4: For an edge to be valid, at least one SNARE pair on the vesicle and target compartment must be active, and have a non-zero entry in the pairing matrix.

\[ \bigwedge\limits_{i,j,q} e_{i,j,q} \supset \bigvee_{k,k^{\prime}} (b_{i,j,q,k} \land a_{j,k^{\prime}} \land p_{k,k^{\prime}}) \, \]

F5: To ensure that fusion respects the graph structure by the edge under consideration, it should not be possible to fuse with any other node.

\[ \bigwedge\limits_{i,j,k} b_{i,j,k} \supset \neg \bigvee_{j \neq j^{\prime}, k^{\prime\prime}} ( a_{j^{\prime},k^{\prime\prime}} \land p_{k,k^{\prime\prime}}) \, \]

\section{Connectivity}
To check the whether that n connected is necessary condition, we remove (drop) n edges from the graph and if it disconnects the graph and we get an assignment we have a n connected satisfying graph. We can go down or up using -C \_ option.  \newline

D0: Only present edges can be dropped.
\[ \bigwedge\limits_{i,j,q} d_{i,j,q} \supset e_{i,j,q}  \]

D1,D2: We are dropping c edges from the graph. exactly c are dropped.
\[ \sum_{i,j,q} d_{i,j,q} == c \]

D3:  Graph becomes disconnected. Ensure that there is no path between some nodes i,j in the underlying undirected graph. Different reachability.
\[ \bigwedge\limits_{i,j} \neg (r^{\prime}_{i,j} \lor r^{\prime}_{j,i})]  \]

D4: New reachability definition for grph connectedness: dReachable. Node i,j are reachable either if there is a direct edge and its not dropped. 
Or there is an node i' such that, there is a direct edge between i,i' which is not dropped and i' and j is dReachable.

\[ \bigwedge\limits_{i,j}  [\bigvee_{q} (e_{i,j,q} \land  \neg d_{i,j,q}) \lor  (\bigvee_{i' != i}  r^{\prime}(i',j) \land  \bigvee_{q} (e_{i,i',q} \land \neg d_{i,i',q}) ] \supset r^{\prime}(i,j)  \]

@@

>>>>

\begin{document}

\title{SAT Solving for Vesicle Traffic Systems in Cells}

\author{to be announced}

\date{\today}

\maketitle

% \vspace*{0.2in}
% % Title must be 250 characters or less.
% \begin{flushleft}
% {\Large
% \textbf\newline{ Vesicle traffic system and constraints.} % Please use 
% }
% \newline
% \\
% \end{flushleft}

\begin{abstract}
\input{abstract}
\end{abstract}

\section{Introduction}
\label{sec:intro}
\input{intro}

% \input{bio}
% \input{conc}

\section{Implementation and Experiments}
\label{sec:experiments}
\input{experiments}

\section{Related Work}
\label{sec:related}
\input{related}

\section{Conclusion}
\label{sec:conclusion}
\input{conclusion}

\bibliographystyle{unsrt}
\bibliography{biblio}

@@

%--------------------- DO NOT ERASE BELOW THIS LINE --------------------------

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "main"
%%% End:
